Tropical Andes plant Lookup Table

### 2023-11-22 code/L1/L1_P2_TropicalAndes_taxcleaning.Rmd
remove NAs and duplicates from all_species list, check lookup table for duplicates, remove duplicates and NA from lookup table, update adding accepted species column to data

```{r}
length(GBIF_species)
length(TRY_species)
length(BIEN_species)
length(GIFT_species)
```
## combine species list together
```{r}
all_species <- c(GBIF_species, TRY_species, BIEN_species, GIFT_species)
all_species <- unique(all_species)
length(all_species)
```
# Remove NAs
```{r}
all_species <- all_species[!is.na(all_species)]
length(all_species)
```

@@

@@

# Disparties
all_species_no.opinion <- subset(all_species_hamonized_results, Taxonomic_status == "No opinion")
nrow(all_species_no.opinion)
all_species_illegitimate <- subset(all_species_hamonized_results, Taxonomic_status == "Illegitimate")
nrow(all_species_illegitimate)
all_species_invalid <- subset(all_species_hamonized_results, Taxonomic_status == "Invalid")
nrow(all_species_invalid)
all_species_other <- subset(all_species_hamonized_results, Taxonomic_status == "")
nrow(all_species_other)
```
# Create species list of names with disparities
```{r}
all_species_disparities <- all_species_hamonized_results %>%
                            filter(Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", "")) %>%
                            select(Name_submitted)
nrow(all_species_disparities)
```

@@

# Save lookup table as csv
```{r}
write.csv(lookup_table, file.path(output_path,"plant_lookup_table.csv"))
```

# Update data files species names using lookup table
## GBIF
```{r}
# filter data to species in lookup_table$Names_submitted
GBIF_occ_subset <- GBIF_occ %>%
  filter(species %in% lookup_table$Name_submitted)
nrow(GBIF_occ_subset)
# add accepted species column from lookup table based on species column
GBIF_occ_subset_harmonized <- merge(GBIF_occ_subset, lookup_table, by.x = "species", by.y = "Name_submitted")
nrow(GBIF_occ_subset_harmonized)
```

## TRY
```{r}

>>>>

```{r}
length(GBIF_species)
length(TRY_species)
length(BIEN_species)
length(GIFT_species)
```
## combine species list together
```{r}
all_species <- c(GBIF_species, TRY_species, BIEN_species, GIFT_species)
all_species <- unique(all_species)
length(all_species)
```

@@

# Remove NAs and duplicates
```{r}
all_species_cleaned <- all_species[!is.na(all_species_cleaned)]
length(all_species_cleaned)
```
```{r}
all_species_cleaned <- unique(all_species_cleaned)
length(all_species_cleaned)
```

@@

# Disparties
all_species_no.opinion <- subset(all_species_hamonized_results, Taxonomic_status == "No opinion")
nrow(all_species_no.opinion)
all_species_illegitimate <- subset(all_species_hamonized_results, Taxonomic_status == "Illegitimate")
nrow(all_species_illegitimate)
all_species_invalid <- subset(all_species_hamonized_results, Taxonomic_status == "Invalid")
nrow(all_species_invalid)
all_species_other <- subset(all_species_hamonized_results, Taxonomic_status == "")
nrow(all_species_other)
all_species_NA <- subset(all_species_hamonized_results, Taxonomic_status == NA)
nrow(all_species_NA)
```
# Create species list of names with disparities
```{r}
all_species_disparities <- all_species_hamonized_results %>%
                            filter(Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", "")) %>%
                            select(Name_submitted)
nrow(all_species_disparities)
```

@@

# Check lookup table for duplicates
```{r}
n_occur <- data.frame(table(lookup_table$Name_submitted))
nrow(n_occur)
duplicate_counts <- n_occur[n_occur$Freq >1,]
nrow(duplicate_counts)
duplicates <- lookup_table[lookup_table$Name_submitted %in% n_occur$Var1[n_occur$Freq > 1],]
nrow(duplicates)
duplicates_pairs <- duplicates[order(duplicates$Name_submitted),]
nrow(duplicates_pairs)
```

# Remove duplicates and accepted names that are NA
```{r}
lookup_table <- lookup_table[!duplicated(lookup_table), ]
nrow(lookup_table)
lookup_table <- subset(lookup_table, !is.na(Accepted_species))
nrow(lookup_table)
```

# Save lookup table as csv
```{r}
write.csv(lookup_table, file.path(output_path,"plant_lookup_table.csv"), row.names = FALSE)
```

# Update data files species names using lookup table
## GBIF
```{r}
# filter data to species in lookup_table$Names_submitted
GBIF_occ_subset <- GBIF_occ %>%
  filter(species %in% lookup_table$Name_submitted)
nrow(GBIF_occ_subset)
# filter lookup_table to species in data
lookup_table_subset <- lookup_table %>%
  filter(Name_submitted %in% GBIF_occ_subset$species)
nrow(lookup_table_subset)

## add accepted species column from lookup table based on species column
GBIF_occ_subset_harmonized <- right_join(lookup_table_subset, GBIF_occ_subset, join_by(Name_submitted == species), keep = TRUE, relationship = "one-to-many")
nrow(GBIF_occ_subset_harmonized)
```

## TRY
```{r}
# filter data to species in lookup_table$Names_submitted
TRY_traits_subset <- TRY_traits %>%
  filter(SpeciesName %in% lookup_table$Name_submitted)
nrow(TRY_traits_subset)
# filter lookup_table to species in data
lookup_table_subset <- lookup_table %>%
  filter(Name_submitted %in% TRY_traits_subset$SpeciesName)
nrow(lookup_table_subset)

## add accepted species column from lookup table based on species column
TRY_traits_subset_harmonized <- right_join(lookup_table_subset, TRY_traits_subset, join_by(Name_submitted == SpeciesName), keep = TRUE, relationship = "one-to-many")
nrow(TRY_traits_subset_harmonized)
```
